timesliced_random_cell_selection

### 2018-11-16 src/biological_data_prep.Rmd
NAO correction :(
nao added to background

nao_yrmt <- nao_yrmt[order(nao_yrmt$year),] #changed amo to nao
nao_yrmt
```

@@

```{r}
data_aea <- merge(x = data_aea, y = nao_yrmt[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
data_aea
```

@@

Note the shifted values are in a new column called amo_sample.Shifted. rename to amo_prev

@@

now extract the WinterAvg.Shifted values to the species dataset

>>>>

nao_yrmt <- nao_yrmt[order(nao_yrmt$year),] 
nao_yrmt
```

@@

```{r}
data_aea <- merge(x = data_aea, y = nao_yrmt[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
head(data_aea)
```

@@

Note the shifted values are in a new column called amo_sample.Shifted. rename to nao_prev

@@

now extract the WinterAvg.Shifted values to the species dataset #####XXXXXXX#####

### 2018-11-16 src/timesliced_random_cell_selection.Rmd
nao added to background

@@

```{r}
background_all <- read.csv("../output/bio/background_all_lonlat_nafo2.csv", header = TRUE)
head(background_all)
```

@@

```{r}
background_all$originalscientificname <- "Mallotus villosus ab"
head(background_all)
```

#NAO and AMO values

>>>>

check for unique years
```{r}
back_yr_uniq <- unique(background_all$year)
back_yr_uniq
```
check for unique months
```{r}
back_mt_uniq <- unique(background_all$month)
back_mt_uniq
```

check number of observations 
```{r}
back_rowcount <- nrow(background_all)
back_rowcount
```

@@

just double-check they are all gone
```{r}
background_all <- read.csv("../output/bio/background_all_lonlat_nafo2.csv", header = TRUE)
bc_nafo2_na <- subset(background_all, is.na(background_all$nafo_zone))
bc_nafo2_na
```
great none!

```{r}
head(background_all)
```
ok so some more crap has been added - the year and month both have _ appended.... change this. Also change glo_unique_cell to cell_id

```{r}
names(background_all)[names(background_all)=="year_"] <- "year"
names(background_all)[names(background_all)=="month_"] <- "month"
names(background_all)[names(background_all)=="glo_unique_cell"] <- "cell_id"
head(background_all)
```

@@

```{r}
background_all$originalscientificname <- "Mallotus villosus ab"
write.csv(background_all, "../output/bio/background_all_lonlat_nafo2_sciname.csv")
head(background_all)
```

# Add AMO Data
The data has been prepared as per the environmental_data_preperation.Rmd, and biological_data_prep.Rmd

So you want..
- AMO value at sampling month/year
- AMO value at previous sampling month/year
- AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)

Load the prepared amo_prev.csv file with has the amo at different year/month steps
```{r}
amo_prev <- read.csv("../output/env/amo_prev.csv", header = TRUE)
head(amo_prev)
```

now match the year and the month in the two dataframes, and populate the background points with amo_sample
```{r}
background_all <- merge(x = background_all, y = amo_prev[ , c("year", "month", "amo_sample")], by = c("year", "month"), all.x = TRUE)
head(background_all)
```

ok next - AMO value at previous sampling month/year 
```{r}
background_all <- merge(x = background_all, y = amo_prev[ , c("year", "month", "amo_prev")], by = c("year", "month"), all.x = TRUE)
head(background_all)
```

and now - AMO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn) (note they are in a different .csv)

```{r}
amo_winter <- read.csv("../output/env/amo_winter.csv")
background_all <- merge(x = background_all, y = amo_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
colnames(background_all)[colnames(background_all)=="WinterAvg.Shifted"] <- "amo_winter" #change the colname
head(background_all)
```

ok and write a csv
```{r}
write.csv(data_aea,"../output/bio/background_all_lonlat_nafo2_sciname_amo.csv", row.names = FALSE)
```

# NAO values
The data has been prepared as per the environmental_data_preperation.Rmd and biological_data_prep.Rmd

So you want..
- NAO value at sampling month/year
- NAO value at previous sampling month/year
- NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn)

Load the prepared nao_prev.csv file with has the amo at different year/month steps
```{r}
nao_prev <- read.csv("../output/env/nao_prev.csv", header = TRUE)
head(nao_prev)
```

now match the year and the month in the two dataframes, and populate the background points with nao_sample

```{r}
background_all <- merge(x = background_all, y = nao_prev[ , c("year", "month", "nao_sample")], by = c("year", "month"), all.x = TRUE)
background_all
```

ok next - NAO value at previous sampling month/year
```{r}
background_all <- merge(x = background_all, y = nao_prev[ , c("year", "month", "nao_prev")], by = c("year", "month"), all.x = TRUE)
background_all
```

and now - NAO phase at previous winter (because the winter values are thought to be a major driver of ocean conditions in the following spring, summer, and autumn). The data is in a different .CSV

```{r}
nao_winter <- read.csv("../output/env/nao_winter.csv")
background_all <- merge(x = background_all, y = nao_winter[ , c("year", "WinterAvg.Shifted")], by = c("year"), all.x = TRUE)
colnames(background_all)[colnames(background_all)=="WinterAvg.Shifted"] <- "nao_winter" #change the colname
background_all
```

and write .csv
```{r}
write.csv(background_all,"../output/bio/background_all_lonlat_nafo2_sciname_amo_nao.csv", row.names = FALSE)
```
